### 2018-07-10 Manuscripts/cog-psych/revision/cogpsych.tex
minor fixes to bib file and intro

Previous work in category learning and generation has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning,smith1995}. Inspired by this work, previous research on the topic of category generation has explored a similar principle: People tend to create new categories that have similar {\em statistical regularities} as previously learned categories \citep{jern2013probabilistic,ward1994structured}. Although this is an important characteristic of generating new categories, it cannot be the only one. Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would be identical to a known category that is representative of the domain (and thus, not new at all).

@@

The second hypothesis for how contrast affects categorization is contrast as {\em representativeness} \citep{kahneman73}. The representativeness heuristic states that people make judgments regarding the probability of an outcome based on how representative it is of the observed evidence. The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by Bayesian inference). For example, the coinflip sequence TTHTH is perceived to be more random than TTTTT (despite both having the same probability of having been generated from a fair coin). Although there is a healthy debate as to whether perceived randomness {\em really} deviates from normative theory \citep{griffiths18,hahn09}, the representativeness heuristic remains one of the main explanations of human judgments \citep{reimers18}. To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}. We do so by extending the state-of-the-art category generation model \citep{jern2013probabilistic} and find that  human category generation is best capture as generating representative, rather than probable, samples.

>>>>

Previous work in category learning and generation has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}. Inspired by this work, previous research on the topic of category generation has explored a similar principle: People tend to create new categories that have similar {\em statistical regularities} as previously learned categories \citep{jern2013probabilistic,ward1994structured}. Although this is an important characteristic of generating new categories, it cannot be the only one. Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would be identical to a known category that is representative of the domain (and thus, not new at all).

@@

The second hypothesis for how contrast might affect categorization is contrast as {\em representativeness}. The representativeness heuristic \citep{kahneman73} states that people make judgments regarding an outcome based on how representative it is of the evidence given in the current context. The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by a straightforward application of Bayes' rule). For example, the coinflip sequence TTHTH is perceived to be more random (due to it being representative of a coinflip sequence) than TTTTT (despite both having the same probability of having been generated from a fair coin). Although there is a healthy debate as to whether perceived randomness {\em really} deviates from normative theory \citep{griffiths18,hahn09}, the representativeness heuristic remains one of the main explanations of human judgments \citep{reimers18}. To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}. We do so by extending the state-of-the-art category generation model \citep{jern2013probabilistic} and find that  human category generation is best capture as generating representative, rather than probable, samples.
